Channel Shaping Using Reconfigurable Intelligent Surfaces: From Diagonal to Beyond

### 2024-02-23 main.tex
:test_tube: Add complexity test
:sparkles: Add complexity test

\usepackage[table]{xcolor}
\usepackage{adjustbox}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bookmark}
\usepackage{booktabs}
\usepackage[makeroom]{cancel}
\usepackage[american]{circuitikz}
\usepackage{cite}
\usepackage{fixmath}
\usepackage[acronym]{glossaries-extra}
\usepackage{hyperref}
\usepackage{import}
\usepackage{mathtools}
\usepackage{microtype}
\usepackage[short]{optidef}
\usepackage{pgfplots}
\usepackage{ragged2e}
\usepackage[subtle]{savetrees}
\usepackage{siunitx}
\usepackage{stfloats}
\usepackage[caption=false,font=footnotesize,subrefformat=parens,labelformat=parens]{subfig}
\usepackage{tabularx}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{epstopdf}
% \usepackage{cleveref}

@@

\begin{section}{\gls{bd}-\gls{ris} Model}
	Consider a \gls{bd}-\gls{ris} aided point-to-point \gls{mimo} system with $N_\mathrm{T}$, $N_\mathrm{S}$, $N_\mathrm{R}$ transmit, scatter, and receive antennas, respectively.
	This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$.
	The \gls{bd}-\gls{ris} is modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that further divides into $G$ individual groups.
	Each group contains $L \triangleq N_\mathrm{S} / G$ elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
	To simplify the analysis, we assume there are no mutual coupling and the in-group connections can be lossless and asymmetric\footnote{While symmetric impedance network is often considered in the literature \cite{Shen2020a,Nerini2023,Santamaria2023,Fang2023,Nerini2023a,Zhou2023,Li2023d,Bartoli2023}, asymmetric passive components (e.g., ring hybrids and branch-line hybrids) may also be reconfigured in real time \cite{Ahn2006}. Asymmetric \gls{bd}-\gls{ris} has been discussed in \cite{Li2023b,Li2023c,Bartoli2023}.}.
	The overall scattering matrix is thus block diagonal $\mathbf{\Theta} = \mathrm{diag}(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G) \in \mathbb{U}^{N_\mathrm{S} \times N_\mathrm{S}}$, where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is a unitary matrix corresponding to group $g \in \mathcal{G} \triangleq \{1,\ldots,G\}$.
	Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$ denote the direct (transmitter-receiver), forward (transmitter-\gls{ris}), and backward (\gls{ris}-receiver) channels, respectively.
	The equivalent channel is
	\begin{equation}
		\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
		\label{eq:channel_equivalent}
	\end{equation}
	where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward channels of \gls{ris} group $g$, respectively.
	\begin{remark}
		\gls{bd}-\gls{ris} reduces to diagonal \gls{ris} and unitary \gls{ris} with group size 1 and $N_\mathrm{S}$, respectively.
	\end{remark}
	\begin{remark}
		Individual forward and backward \gls{csi} are required for \gls{bd}-\gls{ris} designs.
		This is different from diagonal \gls{ris} where estimating their product is usually sufficient.
		% Later we will show the potential benefits from the \gls{csi} overhead.
	\end{remark}
\end{section}

@@

		Inspired by \cite{Abrudan2008,Abrudan2009}, we propose a block-wise \gls{rcg} algorithm along the geodesics on the Lie group of unitary matrices $\mathbb{U}^{L \times L}$.
		It leverages the fact that unitary matrices are closed under multiplication.
		At iteration $r$, the gradient is computed in the Euclidean space and translated to the Riemannian manifold \cite{Absil2009}
		\begin{gather}
			\nabla_{\mathrm{E},g}^{(r)} = \frac{\partial f(\mathbf{\Theta}_g^{(r)})}{\partial \mathbf{\Theta}_g^*},\label{eq:gradient_euclidean}\\
			\nabla_{\mathrm{R},g}^{(r)} = \nabla_{\mathrm{E},g}^{(r)} {\mathbf{\Theta}_g^{(r)}}^\mathsf{H} - \mathbf{\Theta}_g^{(r)} {\nabla_{\mathrm{E},g}^{(r)}}^\mathsf{H}.\label{eq:gradient_riemannian}
		\end{gather}
		The Polak-Ribierre parameter \cite{Polak1969} is approximated as \cite{Abrudan2009}
		\begin{equation}
			\gamma_g^{(r)} = \frac{\mathrm{tr}\bigl((\nabla_{\mathrm{R},g}^{(r)} - \nabla_{\mathrm{R},g}^{(r-1)}) {\nabla_{\mathrm{R},g}^{(r)}}^\mathsf{H}\bigr)}{\mathrm{tr}\bigl(\nabla_{\mathrm{R},g}^{(r-1)} {\nabla_{\mathrm{R},g}^{(r-1)}}^\mathsf{H}\bigr)},
			\label{eq:parameter_cg}
		\end{equation}
		and the conjugate direction is
		\begin{equation}
			\mathbf{D}_g^{(r)} = \nabla_{\mathrm{R},g}^{(r)} + \gamma_g^{(r)} \mathbf{D}_g^{(r-1)}.
			\label{eq:direction_cg}
		\end{equation}
		In the Stiefel manifold, the geodesic emanating from $\mathbf{\Theta}_g^{(r)}$ with velocity $\mathbf{D}_g^{(r)}$ and step size $\mu$ is described compactly by the exponential map \cite{Edelman1998}
		\begin{equation}
			\mathbf{G}_g^{(r)}(\mu) = \exp(\mu \mathbf{D}_g^{(r)}) \mathbf{\Theta}_g^{(r)}.
			\label{eq:geodesic}
		\end{equation}
		An appropriate $\mu^\star$ can be obtained by the Armijo rule \cite{Armijo1966}.\footnote{To double the step size, one only need to square the rotation matrix instead of recomputing the matrix exponential, i.e., $\exp(2 \mu \mathbf{D}_g^{(r)}) = \exp^2(\mu \mathbf{D}_g^{(r)})$.}
		Finally, the scattering matrix is updated along the geodesic as
		\begin{equation}
			\mathbf{\Theta}_g^{(r+1)} = \mathbf{G}_g^{(r)}(\mu^\star).
			\label{eq:update_geodesic}
		\end{equation}
		Algorithm~\ref{ag:rcg} summarizes the proposed block-wise geodesic \gls{rcg} method for smooth maximization problems of asymmetric \gls{bd}-\gls{ris}.
		Convergence to stationary points is guaranteed.

@@

		\begin{proposition}[rank-deficient channel]\label{pp:rank_deficient}
			% If the forward/backward channel is rank-$k$, then a \gls{bd}-\gls{ris} can at most enlarge the $n$-th ($n > k$) channel singular value to the $n-k$-th singular value of $\mathbf{T}$, or suppress the $n$-th ($n < N - k + 1$) channel singular value to the $n$-th singular value of $\mathbf{T}$.
			% If the forward/backward channel is rank-$k$, then a sufficiently large \gls{bd}-\gls{ris} can at most enlarge (resp. suppress) the first (resp. last) $k$ equivalent channel singular values without bounds.
			If the forward or backward channel is rank-$k$, then regardless of the \gls{ris} size and architecture, the $n$-th singular value of the equivalent channel is bounded by
			\begin{align}
				\sigma_n(\mathbf{H}) & \le \sigma_{n-k}(\mathbf{T}), && \text{if } n > k, \label{iq:sv_bound_enlarge}\\
				\sigma_n(\mathbf{H}) & \ge \sigma_n(\mathbf{T}), && \text{if } n < N - k + 1, \label{iq:sv_bound_suppress}
			\end{align}
			where
			% \begin{align}
			% 	\mathbf{T} \mathbf{T}^\mathsf{H} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{V}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}^\mathsf{H}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{F}) = k, \\
			% 	\mathbf{T}^\mathsf{H} \mathbf{T} = \mathbf{H}_\mathrm{D}^\mathsf{H} (\mathbf{I} - \mathbf{U}_\mathrm{F} \mathbf{U}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{B}) = k.
			% \end{align}
			\begin{equation}
				\mathbf{T} \mathbf{T}^\mathsf{H} =
				\begin{cases}
					\mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{V}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}^\mathsf{H}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{F}) = k, \\
					\mathbf{H}_\mathrm{D}^\mathsf{H} (\mathbf{I} - \mathbf{U}_\mathrm{B} \mathbf{U}_\mathrm{B}^\mathsf{H}) \mathbf{H}_\mathrm{D}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{B}) = k,
				\end{cases}
				\label{eq:auxiliary_matrix}
			\end{equation}
			and $\mathbf{V}_\mathrm{F}$ and $\mathbf{U}_\mathrm{B}$ are the right and left compact singular matrices of $\mathbf{H}_\mathrm{F}$ and $\mathbf{H}_\mathrm{B}$, respectively.
			% where $\mathbf{T} \mathbf{T}^\mathsf{H} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{V}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}$ if the forward channel is rank-$k$ and $\mathbf{T}^\mathsf{H} \mathbf{T} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{U}_\mathrm{F} \mathbf{U}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}$ if the backward channel is rank-$k$.
		\end{proposition}
		\begin{proof}
			Please refer to Appendix~\ref{ap:rank_deficient}.
		\end{proof}

@@

@@

\begin{appendix}
	\begin{subsection}{Proof of Lemma \ref{lm:pareto_gradient}}\label{ap:pareto_gradient}
		Let $\mathbf{H} = \sum_n \mathbf{u}_n \sigma_n \mathbf{v}_n^\mathsf{H}$ be the compact \gls{svd} of the equivalent channel.
		Since the singular vectors are orthonormal, the $n$-th singular value can be expressed as
		\begin{equation}
			\sigma_n = \mathbf{u}_n^\mathsf{H} \mathbf{H} \mathbf{v}_n = \mathbf{u}_n^\mathsf{T} \mathbf{H}^* \mathbf{v}_n^*,
		\end{equation}
		whose differential w.r.t. $\mathbf{\Theta}_g^*$ is
		\begin{align*}
			\partial \sigma_n
			& = \partial \mathbf{u}_n^\mathsf{T} \underbrace{\mathbf{H}^* \mathbf{v}_n^*}_{\sum_m \mathbf{u}_m^* \sigma_m \mathbf{v}_m^\mathsf{T} \mathbf{v}_n} + \mathbf{u}_n^\mathsf{T} \cdot \partial \mathbf{H}^* \cdot \mathbf{v}_n^* + \underbrace{\mathbf{u}_n^\mathsf{T} \mathbf{H}^*}_{\mathbf{u}_n^\mathsf{T} \sum_m \mathbf{u}_m^* \sigma_m \mathbf{v}_m^\mathsf{T}} \partial \mathbf{v}_n^*\\
			& = \underbrace{\partial \mathbf{u}_n^\mathsf{T} \mathbf{u}_n^*}_{\partial 1 = 0} \cdot \sigma_n + \mathbf{u}_n^\mathsf{T} \cdot \partial \mathbf{H}^* \cdot \mathbf{v}_n^* + \sigma_n \cdot \underbrace{\mathbf{v}_n^\mathsf{T} \partial \mathbf{v}_n^*}_{\partial 1 = 0}\\
			& = \mathbf{u}_n^\mathsf{T} \mathbf{H}_{\mathrm{B},g}^* \cdot \partial \mathbf{\Theta}_g^* \cdot \mathbf{H}_{\mathrm{F},g}^* \mathbf{v}_n^*\\
			& = \mathrm{tr}(\mathbf{H}_{\mathrm{F},g}^* \mathbf{v}_n^*\mathbf{u}_n^\mathsf{T} \mathbf{H}_{\mathrm{B},g}^* \cdot \partial \mathbf{\Theta}_g^*).
		\end{align*}
		According to \cite{Hjorungnes2007}, the corresponding complex derivative is
		\begin{equation}
			\frac{\partial \sigma_n}{\partial \mathbf{\Theta}_g^*} = \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \mathbf{u}_n \mathbf{v}_n^\mathsf{H} \mathbf{H}_{\mathrm{F},g}^\mathsf{H}.
			\label{eq:derivative_sv}
		\end{equation}
		A linear combination of \eqref{eq:derivative_sv} yields \eqref{eq:pareto_gradient}.
	\end{subsection}

>>>>

\usepackage[table]{xcolor}
\usepackage{adjustbox}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bookmark}
\usepackage{booktabs}
\usepackage[makeroom]{cancel}
\usepackage[american]{circuitikz}
\usepackage{cite}
\usepackage{fixmath}
\usepackage[acronym]{glossaries-extra}
\usepackage{hyperref}
\usepackage{import}
\usepackage{mathtools}
\usepackage{microtype}
\usepackage[short]{optidef}
\usepackage{pgfplots}
\usepackage{ragged2e}
\usepackage[subtle]{savetrees}
\usepackage{siunitx}
\usepackage{stfloats}
\usepackage[caption=false,font=footnotesize,subrefformat=parens,labelformat=parens]{subfig}
\usepackage{tabularx}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{multirow}
\usepackage{booktabs}
% \usepackage{cleveref}

@@

\begin{section}{\gls{bd}-\gls{ris} Model}
	Consider a \gls{bd}-\gls{ris} aided point-to-point \gls{mimo} system with $N_\mathrm{T}$, $N_\mathrm{S}$, $N_\mathrm{R}$ transmit, scatter, and receive antennas, respectively.
	This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$ in the following context.
	The \gls{bd}-\gls{ris} can be modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that divides into $G$ individual groups, each containing $L \triangleq N_\mathrm{S} / G$ elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
	To simplify the analysis, we assume a lossless asymmetric network \footnote{While symmetric \gls{bd}-\gls{ris} (involving capacitors and inductors) is often considered in the literature \cite{Shen2020a,Nerini2023,Santamaria2023,Fang2023,Nerini2023a,Zhou2023,Li2023d,Bartoli2023}, asymmetric \gls{bd}-\gls{ris} can be built over reconfigurable asymmetric passive components (e.g., ring hybrids and branch-line hybrids) \cite{Ahn2006}.} without mutual coupling between scatter elements, as previously considered in \cite{Li2023b,Li2023c,Bartoli2023}.
	The overall scattering matrix of the \gls{bd}-\gls{ris} is
	\begin{equation}
		\mathbf{\Theta} = \mathrm{diag}(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),
	\end{equation}
	where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is a unitary matrix describing the scatter response of group $g$.
	Note that diagonal (single-connected) and unitary (fully-connected) \gls{ris} can be regarded as its extreme cases with group size 1 and $N_\mathrm{S}$, respectively.
	The potential physical architectures of \gls{bd}-\gls{ris} are shown in \cite[Fig. 3]{Shen2020a}, \cite[Fig. 5]{Li2023c}, and \cite[Fig. 2]{Nerini2023a}, where the antenna radiation patterns and circuit network structures are taken into account.
	% Note that \gls{bd}-\gls{ris} reduces to diagonal \gls{ris} and unitary \gls{ris} with group size equal to 1 and $N_\mathrm{S}$, respectively.
	% Note that \gls{bd}-\gls{ris} reduces to diagonal \gls{ris} and unitary \gls{ris} with group size equal to 1 and $N_\mathrm{S}$, respectively.

	Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$ denote the direct (transmitter-receiver), forward (transmitter-\gls{ris}), and backward (\gls{ris}-receiver) channels, respectively.
	The equivalent channel is
	\begin{equation}
		\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
		\label{eq:channel_equivalent}
	\end{equation}
	where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward channels of \gls{ris} group $g$, respectively.
	\begin{remark}
		\gls{bd}-\gls{ris} reduces to diagonal \gls{ris} and unitary \gls{ris} with group size $L$ 1 and $N_\mathrm{S}$, respectively.
	\end{remark}
	\begin{remark}
		Individual forward and backward \gls{csi} are required for \gls{bd}-\gls{ris} designs.
		This is different from diagonal \gls{ris} where estimating their product is usually sufficient.
		% Later we will show the potential benefits from the \gls{csi} overhead.
	\end{remark}
\end{section}

@@

		Inspired by \cite{Abrudan2008,Abrudan2009}, we propose a block-wise \gls{rcg} algorithm along the geodesics on the Lie group of unitary matrices $\mathbb{U}^{L \times L}$.
		It leverages the fact that unitary matrices are closed under multiplication.
		At iteration $r$, the gradient is computed in the Euclidean space and translated to the Riemannian manifold \cite{Absil2009}
		\begin{gather}
			\nabla_{\mathrm{E},g}^{(r)} = \frac{\partial f(\mathbf{\Theta}_g^{(r)})}{\partial \mathbf{\Theta}_g^*},\label{eq:gradient_euclidean}\\
			\nabla_{\mathrm{R},g}^{(r)} = \nabla_{\mathrm{E},g}^{(r)} {\mathbf{\Theta}_g^{(r)}}^\mathsf{H} - \mathbf{\Theta}_g^{(r)} {\nabla_{\mathrm{E},g}^{(r)}}^\mathsf{H}.\label{eq:gradient_riemannian} % $\mathcal{O}(N_\mathrm{S}^3)$
		\end{gather}
		The Polak-Ribierre parameter \cite{Polak1969} is approximated as \cite{Abrudan2009}
		\begin{equation}
			\gamma_g^{(r)} = \frac{\mathrm{tr}\bigl((\nabla_{\mathrm{R},g}^{(r)} - \nabla_{\mathrm{R},g}^{(r-1)}) {\nabla_{\mathrm{R},g}^{(r)}}^\mathsf{H}\bigr)}{\mathrm{tr}\bigl(\nabla_{\mathrm{R},g}^{(r-1)} {\nabla_{\mathrm{R},g}^{(r-1)}}^\mathsf{H}\bigr)}, % $\mathcal{O}(2 N_\mathrm{S}^3 + N_\mathrm{S}^2 + 2 N_\mathrm{S})$
			\label{eq:parameter_cg}
		\end{equation}
		and the conjugate direction is
		\begin{equation}
			\mathbf{D}_g^{(r)} = \nabla_{\mathrm{R},g}^{(r)} + \gamma_g^{(r)} \mathbf{D}_g^{(r-1)}. % $\mathcal{O}(N_\mathrm{S}^2)$
			\label{eq:direction_cg}
		\end{equation}
		In the Stiefel manifold, the geodesic emanating from $\mathbf{\Theta}_g^{(r)}$ with velocity $\mathbf{D}_g^{(r)}$ and step size $\mu$ is described compactly by the exponential map \cite{Edelman1998}
		\begin{equation}
			\mathbf{G}_g^{(r)}(\mu) = \exp(\mu \mathbf{D}_g^{(r)}) \mathbf{\Theta}_g^{(r)}. % $\mathcal{O}(N_\mathrm{S}^3)$
			\label{eq:geodesic}
		\end{equation}
		An appropriate $\mu^\star$ can be obtained by the Armijo rule \cite{Armijo1966}.\footnote{To double the step size, one only need to square the rotation matrix instead of recomputing the matrix exponential, i.e., $\exp(2 \mu \mathbf{D}_g^{(r)}) = \exp^2(\mu \mathbf{D}_g^{(r)})$.}
		Finally, the scattering matrix is updated along the geodesic as
		\begin{equation}
			\mathbf{\Theta}_g^{(r+1)} = \mathbf{G}_g^{(r)}(\mu^\star).
			\label{eq:update_geodesic}
		\end{equation}
		Algorithm~\ref{ag:rcg} summarizes the proposed block-wise geodesic \gls{rcg} method for smooth maximization problems of asymmetric \gls{bd}-\gls{ris}.
		Convergence to stationary points is guaranteed.

@@

		\begin{proposition}[rank-deficient channel]\label{pp:rank_deficient}
			% If the forward/backward channel is rank-$k$, then a \gls{bd}-\gls{ris} can at most enlarge the $n$-th ($n > k$) channel singular value to the $n-k$-th singular value of $\mathbf{T}$, or suppress the $n$-th ($n < N - k + 1$) channel singular value to the $n$-th singular value of $\mathbf{T}$.
			% If the forward/backward channel is rank-$k$, then a sufficiently large \gls{bd}-\gls{ris} can at most enlarge (resp. suppress) the first (resp. last) $k$ equivalent channel singular values without bounds.
			If the forward or backward channel is rank-$k$, then regardless of the \gls{ris} size and architecture, the $n$-th singular value of the equivalent channel is bounded by
			\begin{align}
				\sigma_n(\mathbf{H}) & \le \sigma_{n-k}(\mathbf{T}), &  & \text{if } n > k, \label{iq:sv_bound_enlarge}          \\
				\sigma_n(\mathbf{H}) & \ge \sigma_n(\mathbf{T}),     &  & \text{if } n < N - k + 1, \label{iq:sv_bound_suppress}
			\end{align}
			where
			% \begin{align}
			% 	\mathbf{T} \mathbf{T}^\mathsf{H} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{V}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}^\mathsf{H}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{F}) = k, \\
			% 	\mathbf{T}^\mathsf{H} \mathbf{T} = \mathbf{H}_\mathrm{D}^\mathsf{H} (\mathbf{I} - \mathbf{U}_\mathrm{F} \mathbf{U}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{B}) = k.
			% \end{align}
			\begin{equation}
				\mathbf{T} \mathbf{T}^\mathsf{H} =
				\begin{cases}
					\mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{V}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}^\mathsf{H}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{F}) = k, \\
					\mathbf{H}_\mathrm{D}^\mathsf{H} (\mathbf{I} - \mathbf{U}_\mathrm{B} \mathbf{U}_\mathrm{B}^\mathsf{H}) \mathbf{H}_\mathrm{D}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{B}) = k,
				\end{cases}
				\label{eq:auxiliary_matrix}
			\end{equation}
			and $\mathbf{V}_\mathrm{F}$ and $\mathbf{U}_\mathrm{B}$ are the right and left compact singular matrices of $\mathbf{H}_\mathrm{F}$ and $\mathbf{H}_\mathrm{B}$, respectively.
			% where $\mathbf{T} \mathbf{T}^\mathsf{H} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{V}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}$ if the forward channel is rank-$k$ and $\mathbf{T}^\mathsf{H} \mathbf{T} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{U}_\mathrm{F} \mathbf{U}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}$ if the backward channel is rank-$k$.
		\end{proposition}
		\begin{proof}
			Please refer to Appendix~\ref{ap:rank_deficient}.
		\end{proof}

@@

	\begin{table*}[!t]
		\caption{Average Performance of \gls{bd}-\gls{ris} Designs}
		\label{tb:complexity_test}
		\centering
		\begin{tabular}{ccccccc}
			\toprule
			\multirow{2}{*}{\gls{rcg} path} & \multicolumn{3}{c}{$N_\mathrm{S}=16$} & \multicolumn{3}{c}{$N_\mathrm{S}=256$}                                                             \\ \cmidrule(lr){2-4} \cmidrule(lr){5-7}
			                                & Objective                             & Iterations                             & Time [s]        & Objective       & Iterations & Time [s] \\ \midrule
			Geodesic                        & $\num{4.36e-3}$                       & 11.61                                  & $\num{2.04e-2}$ & $\num{1.16e-2}$ & 25.78      & 3.21     \\
			Non-geodesic                    & $\num{4.17e-3}$                       & 169.50                                 & $\num{1.42e-1}$ & $\num{8.87e-3}$ & 278.08     & 27.81    \\ \bottomrule
		\end{tabular}
	\end{table*}

@@

\begin{appendix}
	\begin{subsection}{Proof of Lemma \ref{lm:pareto_gradient}}\label{ap:pareto_gradient}
		Let $\mathbf{H} = \sum_n \mathbf{u}_n \sigma_n \mathbf{v}_n^\mathsf{H}$ be the compact \gls{svd} of the equivalent channel.
		Since the singular vectors are orthonormal, the $n$-th singular value can be expressed as
		\begin{equation}
			\sigma_n = \mathbf{u}_n^\mathsf{H} \mathbf{H} \mathbf{v}_n = \mathbf{u}_n^\mathsf{T} \mathbf{H}^* \mathbf{v}_n^*,
		\end{equation}
		whose differential w.r.t. $\mathbf{\Theta}_g^*$ is
		\begin{align*}
			\partial \sigma_n
			 & = \partial \mathbf{u}_n^\mathsf{T} \underbrace{\mathbf{H}^* \mathbf{v}_n^*}_{\sum_m \mathbf{u}_m^* \sigma_m \mathbf{v}_m^\mathsf{T} \mathbf{v}_n} + \mathbf{u}_n^\mathsf{T} \cdot \partial \mathbf{H}^* \cdot \mathbf{v}_n^* + \underbrace{\mathbf{u}_n^\mathsf{T} \mathbf{H}^*}_{\mathbf{u}_n^\mathsf{T} \sum_m \mathbf{u}_m^* \sigma_m \mathbf{v}_m^\mathsf{T}} \partial \mathbf{v}_n^* \\
			 & = \underbrace{\partial \mathbf{u}_n^\mathsf{T} \mathbf{u}_n^*}_{\partial 1 = 0} \cdot \sigma_n + \mathbf{u}_n^\mathsf{T} \cdot \partial \mathbf{H}^* \cdot \mathbf{v}_n^* + \sigma_n \cdot \underbrace{\mathbf{v}_n^\mathsf{T} \partial \mathbf{v}_n^*}_{\partial 1 = 0}                                                                                                                  \\
			 & = \mathbf{u}_n^\mathsf{T} \mathbf{H}_{\mathrm{B},g}^* \cdot \partial \mathbf{\Theta}_g^* \cdot \mathbf{H}_{\mathrm{F},g}^* \mathbf{v}_n^*                                                                                                                                                                                                                                                 \\
			 & = \mathrm{tr}(\mathbf{H}_{\mathrm{F},g}^* \mathbf{v}_n^*\mathbf{u}_n^\mathsf{T} \mathbf{H}_{\mathrm{B},g}^* \cdot \partial \mathbf{\Theta}_g^*).
		\end{align*}
		According to \cite{Hjorungnes2007}, the corresponding complex derivative is
		\begin{equation}
			\frac{\partial \sigma_n}{\partial \mathbf{\Theta}_g^*} = \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \mathbf{u}_n \mathbf{v}_n^\mathsf{H} \mathbf{H}_{\mathrm{F},g}^\mathsf{H}.
			\label{eq:derivative_sv}
		\end{equation}
		A linear combination of \eqref{eq:derivative_sv} yields \eqref{eq:pareto_gradient}.
	\end{subsection}
